### 2024-02-27 writing/paper.Rmd
organizing and adding first paper structure draft

---
title: "Untitled"
author: "Bennour Mohamed Hsin"
date: "2024-02-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

  importance of education for properity and sustainable development. education and the motor of innovation.

  the literature trends in assessing school systems (what is important for school systems, make the right things vs making things right).

  using efficiency analysis as a methodology to analyze school systems as a proposition to solve the problem.

## Methodology

Description of the data, sources and preparation

descriptive summary of the data

description of the DEA model and its parameters and specifications, its use on similar data and the expected results

## Results

results of the efficiency, multipliers and the intensity weights

all the plots

## Discussion

limitations of the dea methodology and the specifications.

how these results can have policy implications

## Conclusion

summarize.

## References