### 2024-03-22 presentation/presentation_bennour.Rmd
adding presentation

---
title: 'Visegrad Countries and Education Challenges : An Efficiency Analysis'
author: "Bennour Mohamed Hsin - University of Pécs - Faculty of Business and Economics"
date: "2024-03-22"
output:
  slidy_presentation: default
  ioslides_presentation: default
  beamer_presentation: default
bibliography: refs.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## AGENDA

- Context

- Methodology

- Results

- Possible extensions

- Conclusion

## CONTEXT

- Capitalism requires a push for efficiency
  - either tby maximizing output or minimizing input
  - Either ways innovation is key

- Knowledge is the engine that drives innovation (among other things) [@cunha2006].

- Knowledge capital increasing is the result of the investment efforts in Human Capital (**AND EDUCATION**) [@hanushek2021]

- The importance of Education in increasing the levels of knowledge and thus the capacity to innovate in the long run, affecting gorwth and potentially development and wellfare [@hanushek2020_a].

## --

- Assessing Education performance is necessary to drive policies in a larger context of innovation policy.

- [@worthington2001] assumes that education institutions conduct a basic transformation of inputs into outputs, and this transformation can be proxied with a production function

- Assessing the efficiency based on input output approach is plausible in this context

## --

- What are the drivers of efficiency for the V4 schools?

- What are the environmental factors that affect this efficiency?

## METHODOLOGY

- Methodology for evaluating educational institutions includes DEA (Data Envelopment Analysis [@charnes1978]), combined with other methods like Bootstrapping, regression (OLS, Tobit, or truncated), or the Malmquist index.

- Focus on school-level analysis, treating educational institutions as units that transform inputs into outputs.

- Schools are assumed to be influenced by various internal and external factors.

##

- Investigation involves two-stage analysis using 2022 PISA survey data for schools in the Visegrad countries:
  - First stage: Estimate efficiency of schools in each country (output oriented).
  - Second stage: Double bootstrap procedure with truncated regression [@simar2007] to assess environmental factors' influence on school efficiency.

## DATA

| Category                | Determinants                                                                                                           |
|-------------------------|------------------------------------------------------------------------------------------------------------------------|
| Inputs                  | STRATIO (Students to Teacher Ratio), CPERIODS (Class Periods per Week)                                                 |
| Outputs                 | MATH, READING, SCIENCE (Average Test Scores)                                                                           |
| Environmental Variables | STUBI (Student Behavioral Issues), TEABI (Teacher Behavioral Issues), STAFI (Staff Issues), RESSI (Resources), COMP (Competition), LOC (Location), SCSIZE (School Size), TOTAT (Total Number of Teachers) |

## RESULTS: STAGE 1

- First Stage DEA Estimation:
  - Poland is the most efficient country (score: 1.183).
  - Slovakia follows with a score of 1.226.
  - Hungary has a score of 1.253 with the largest efficiency variation (0.172).
  - Czech Republic is the least efficient (score: 1.269).

## --

- Distribution of Schools Across Efficiency Quartiles:
  - Poland has the highest percentage of fully efficient schools (5.46%).
  - Hungary follows with 3.54%, Slovakia with 3.32%, and Czech Republic with 1.92%.
  - Poland also has the highest percentage of highly inefficient schools (24.79%), followed by Czech Republic (23.98%).

## --

- Influence of School Subjects on Efficiency:
  - Science influences efficiency most in Polish and Czech schools (55.56% and 48.62% respectively).
  - Mathematics is most influential in Hungarian (39.81%) and Slovakian schools (73.98%).

## RESULTS: SECOND STAGE

  - Positive and significant baseline efficiencies across all countries, with Hungary having the highest.
  - Student behavioral issues (STUBI) positively impact efficiency in all countries, with the highest impact in Hungary.
  - Teacher behavior issues (TEABI) negatively affect efficiency, particularly in Hungary.
  - In Poland, resource issues (RESSI) significantly reduce school efficiency, unlike in other countries.
  - Czech Republic sees a positive effect on efficiency from staff issues (STAFI).

## --

  - School competition (COMP) improves efficiency significantly only in Czech Republic and Poland.
  - School size (SCSIZE) shows a significant positive impact only in Czech Republic.
  - Schools in rural areas (LOC) are more efficient than urban ones, especially in Hungary.
  - Number of teachers (TOTAT) negatively impacts efficiency in Czech Republic and Hungary; not significant in Poland and Slovakia.

## Implications

Here's a table summarizing the recommendations for each country:

| Country   | Recommendation 1                                      | Recommendation 2                                             | Recommendation 3                                              | Recommendation 4                                            |
|-----------|-------------------------------------------------------|--------------------------------------------------------------|---------------------------------------------------------------|-------------------------------------------------------------|
| **Poland**    | Enhance Resource Allocation                           | Expand Successful Schools                                    | Address Inefficiencies in Large Schools                       | Foster Healthy Competition                                  |
| **Slovakia**  | Focus on Teacher Quality                              | Leverage Efficient Small Schools                             | Enhance Student Behavior Programs                            | Address Staff Issues                                        |
| **Hungary**   | Improve Student Behavioral Interventions             | Enhance Teacher Support and Development                      | Optimize School Size                                         | Promote Rural School Efficiency                             |
| **Czech Republic** | Invest in Staff Development                           | Optimize Teacher Numbers                                     | Encourage School Competition                                 | Improve Resource Efficiency                                 |

## POTENTIAL ISSUES and POSSIBLE EXTENSIONS

- Aggregation on different levels might be problematic

- Separability test was not conducted

- Directional Distance Function might be more appropriate than DEA in this case

- Adding time dimension can increase the importance of the study

# Thank you for your attention